Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2018-01-18 paper-ieee-trans.tex
Rephrase and complete introduction

\IEEEPARstart{D}uring the last 15 years the provision of a European e-Infrastructure 
has been driven by the needs of user communities to access large scale computing and 
data analysis infrastructure to support their research needs as part of their 
pan-European research collaborations. One of the most known examples is the Worldwide 
Large Hadron Collider Computing Grid (WLCG) global computing infrastructure, powered 
by the Particle Physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing the huge quantity of data coming from the Large 
Hadron Collider particle accelerator. As a consequence a number of important projects 
have been run in Europe, supported by the European Commission funding of various 
research programms (Framework Programms 5th, 6th, 7th, Horizon 2020) in order to
develop distributed solutions that provided a seamless and transparent access to those 
resources, available for the researchers.

Major developments were done throughout the years, mainly focused on providing new
functionalities, leaving secondly the adoption of software engineering methodologies.
The software delivered was then less tested, resulting in constant
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e-Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.

@@

Analyzing the European Commission-funded software projects in the last decade
we have observed a continuous increase of the prominence and robustness of
testing and validation procedures. One of many reasons is related to the
evolution of the software engineering techniques (e.g. creation of the Agile
Manifesto \cite{agile-manifesto}, rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the ICT field with virtualization and
container technologies. Another one is the emergence of automation and
event-response tools.

In this paper, we introduce a set of European Commission (EC)-funded projects
that goes from 2001 up to now, in which the authors of this summary were involved.
For each project, we describe the challenges arised, the methodologies used and
the achievements towards software reliability. The reminder of this paper is
organized as follows. Section \ref{sec:ev} details the evolution of research on
software reliability in EC-projects.  Section \ref{sec:ntsr} provides
information about new research trends in software reliability. Finally,
Section \ref{sec:con} concludes.

>>>>

\IEEEPARstart{D}uring the last 15 years the provision of an European e-Infrastructure,
that supports an unified access to large-scale computing and intense data analysis
applications, has been driven by the needs of scientific communities as part of their
pan-European research collaborations. One of the well-known examples is the Worldwide 
Large Hadron Collider Computing Grid (WLCG) global computing infrastructure, powered 
by the Particle Physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing the huge quantity of data coming from the Large 
Hadron Collider particle accelerator. A significant number of innovation projects and
technology transfer activities -- embraced by the 5th, 6th and 7th Framework Programmes,
and ultimately by the Horizon 2020 programme \footnote{https://ec.europa.eu/programmes/horizon2020/} -- are being held to coordinate and 
develop distributed solutions that provide a seamless and transparent access to those
computing and data resources, available for the researchers.

Major developments were done throughout the years, mainly focused on providing new
functionalities, leaving aside at early times the adoption of software engineering methodologies.
The software delivered was then less tested, resulting in constant
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e-Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.

@@

Analyzing the European Commission (EC)-funded software projects in the last decade
we have observed a continuous increase of the prominence and robustness of
testing and validation procedures. One of many reasons is related to the
evolution of the software engineering methodologies (e.g. creation of the Agile
Manifesto \cite{agile-manifesto}, rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the information and communications technology (ICT)
field such as virtualization, enabling technology for Cloud computing and 
operating--system--level virtualization known as containerization. All of which
complemented by the emergence of automation and event-response tools.

In this paper, we introduce a set of EC-funded projects
that go from 2001 up to now, in which the authors of this summary were involved.
For each project, we describe the challenges arised, the methodologies used and
the achievements towards software reliability. The reminder of this paper is
organized as follows. Section \ref{sec:ev} details the evolution of research on
software reliability in EC-projects. Section \ref{sec:ntsr} provides
information about new research trends in software reliability based upon the 
experiences from the recent INDIGO-DataCloud project. Finally, Section \ref{sec:con} concludes.
